Embedding Dimension of Polyhedral Losses

### 2020-06-19 colt-20/embed-dim.tex
camera edits

\documentclass[final]{colt2020} % Anonymized submission
% \documentclass{colt2020} % Include author names

@@

\usepackage{mathtools, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
	linkcolor=darkblue,urlcolor=darkblue,
	anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}

@@

Optimizing such discrete losses over a dataset (i.e. empirical risk minimization), however, is typically computationally hard.
One therefore generally resorts to optimizing a computationally nice \emph{surrogate loss} $L$.
The key requirement is \emph{calibration}, namely that optimizing surrogate loss allows one to recover the discrete prediction that optimizes the original $\ell$-loss.
Calibrated surrogate losses yields desirable statistical guarantees such as the existence of excess risk bounds given by consistency~\citep{tewari2007consistency}.
%In the finite outcome setting, we know that calibration implies consistency, so we study the design of calibrated, convex losses-- specifically polyhedral (piecewise linear and convex) losses.

The question therefore becomes: Given a discrete loss $\ell$, how do we design such calibrated surrogates $L$? 
In this paper we study \emph{polyhedral} surrogates, meaning piecewise-linear convex functions over prediction space $\reals^d$.
\citet{finocchiaro2019embedding} show that polyhedral surrogates are in a strong sense equivalent to an \emph{embedding}: % \emph{embedding} the discrete loss $\ell$ into $\reals^d$ by 
mapping each discrete prediction $r$ to some point $\varphi(r) \in \reals^d$, which optimizes $L$-loss if and only if $r$ optimizes $\ell$-loss.
While the authors show that every discrete loss $\ell$ can be so embedded, in the worst case the dimension required is $d=n-1$ where $n$ is the number of possible observations.
A surrogate dimension $d$ significantly below $n$, such as $O(\log n)$ for classification with an abstain option~\citep{ramaswamy2018consistent}, can lead to faster downstream optimization and computation, an effect that grows with $n$; this motivates us to understand when this dimension can be low.

@@

%In this work, we study \emph{embedding dimension}, which is the minimum dimension of input to a polyhedral surrogate that is calibrated with respect to the given discrete loss.
%Moreover, we use this tool to characterize \emph{how} embeddable a given discrete loss is, showing necessary conditions for embedding a given loss into $d$ dimensions.
In order to concisely discuss embedding dimension, we appeal to notation and terminology from the field of property elicitation (\citep{savage1971elicitation, osband1985information-eliciting, gneiting2007strictly, lambert2009eliciting, lambert2018elicitation}), relating it to the language of calibrated surrogates as needed.

@@

\begin{definition}[Embedding dimension]
  We say a discrete loss $\ell: \R \to \nonnegreals^{\Y}$ is \emph{$d$-embeddable} if it can be embedded into $d$ dimensions, i.e. if there exists a calibrated polyhedral surrogate $L: \reals^d \to \nonnegreals^{\Y}$. \discuss{Jessie- Can we revisit this definition post i.e.?}
  The \emph{embedding dimension} of $\ell$ is the smallest $d$ such that $\ell$ is $d$-embeddable.
\end{definition}

@@

We summarize the above results in the following theorem, together with one additional result, whose proof is in Appendix~\ref{app:1-dim}: if $\ell$ has any calibrated convex surrogate at all, it must have a polyhedral one.
(See Conjecture~\ref{conj:embed-equals-ccdim} regarding higher dimensional generalizations.)
\begin{theorem} \label{thm:1d-tfae}
  Let $\ell$ be a discrete loss eliciting a finite property $\gamma$.
  The following are equivalent: (1) $\gamma$ is orderable; (2) the intersection graph of $\gamma$ is a path; (3) the two conditions of Theorem \ref{thm:construct-1d-loss} are satisfied; (4) $\ell$ is $1$-embeddable; (5) $\ell$ has some polyhedral calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$; (6) $\ell$ has some convex calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
\end{theorem}
\jessiet{If I understand correctly, different orderings of the reports may yield different embeddings, and these different embeddings should elicit different properties.  How to we reconcile this in the theorem statement?  Or where am I wrong?}

@@

Recall that if $\ell: \R \to \nonnegreals^{\Y}$ is embedded by $L: \reals^d \to \nonnegreals^{\Y}$, then each $r \in \R$ is embedded at some point $\varphi(r) \in \reals^d$.
In particular, $\varphi(r)$ must minimize $\inprod{p}{L(\cdot)}$ if and only if $r$ minimizes $\inprod{p}{\ell(\cdot)}$.
The key to our approach is to study as first-class objects the sets of all subgradients\footnote{Recall that a subgradient of e.g. the convex function $L(\cdot)_y: \reals^d \to \reals$ at a point $u$ is a vector $v \in \reals^d$ such that $L(u')_y \geq L(u)_y + \inprod{v}{u'-u}$ for all $u'$.} of $L$ at these embedding points.
The question of whether a calibrated polyhedral surrogate exists in $d$ dimensions essentially reduces to conditions on these sets alone.
\raft{The ``essentially'' is because we need loss values to match.  It's not clear if that comes for free, i.e., if you embed the correct property, can you always modify the loss to match $\ell$?}
In particular, we use the fact that a convex function is minimized at $u$ if and only if $\vec{0}$ is in its subgradient set at $u$.
Therefore, we consider collections of sets $T^r_y$, which intuitively aspire to be the subgradient sets of a calibrated polyhedral surrogate $L(\cdot)_y$ at $\varphi(r)$, denoted $\partial L(\varphi(r))_y$.
Throughout, we often take $r$ as implicit and suppress it from our notation for ease of exposition.
Note that if $L(\cdot)_y$ is a polyhedral function on $\reals^d$, then all of its subgradient sets are (bounded) closed polytopes~\citep{rockafellar1997convex}.

@@

  \vspace{-3.2em}
    \begin{align}
      V X^j                     &\leq B    & \text{(pointwise, $\forall j \in [\ell]$)}  \label{eqn:qp-constr-1} \\
      \sum_{y=1}^n p^j_y x^j_y &= \vec 0  & \text{($\forall j \in [\ell]$)}    \label{eqn:qp-constr-2}
    \end{align}
\end{definition}

@@

%% Bo: I think we can get away with no summary at the end.
%\paragraph{Conclusion.}
%\jessie{Not a fan, but a placeholder if we get too close to deadline.}
%In this paper, we present the notion of \emph{embedding dimension} in order to measure the efficiency of a surrogate loss function.
%While similar notions of surrogate efficiency have been studied before, we propose a characterization in $1$ dimension that generalizes to the construction of any calibrated convex surrogate.
%In higher dimensions, our characterization can only be applied to understanding the minimal dimension of a calibrated \emph{polyhedral} surrogate, although we conjecture that embedding dimension is equal to convex calibration dimension in general.
%In order to demonstrate our characterizations, we show an improved lower bound from previous work on the convex calibration dimension of $\abstain{\alpha}$, and a new lower bound on embedding dimension of $\abstain{1/2}$.
%
\newpage
% Acknowledgments---Will not appear in anonymized version
\acks{Jessie Finocchiaro was funded by NSF Graduate Research Fellowship under grant \# DGE 1650115. \raf{Others...?} \bo{Others...?}}

>>>>

\documentclass[final]{colt2020} % Anonymized submission
% \renewcommand{\includegraphics}[1][1]{}

% \documentclass{colt2020} % Include author names

@@

\usepackage{mathtools, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
	linkcolor=darkblue,urlcolor=darkblue,
	anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}

@@

Optimizing such discrete losses over a dataset (i.e. empirical risk minimization), however, is typically computationally hard.
One therefore generally resorts to optimizing a computationally nice \emph{surrogate loss} $L$.
The key requirement is \emph{calibration}, namely that optimizing surrogate loss allows one to recover the discrete prediction that optimizes the original $\ell$-loss.
Calibrated surrogate losses yields desirable statistical guarantees such as consistency and excess risk bounds~\citep{tewari2007consistency}.
%In the finite outcome setting, we know that calibration implies consistency, so we study the design of calibrated, convex losses-- specifically polyhedral (piecewise linear and convex) losses.

The question therefore becomes: Given a discrete loss $\ell$, how do we design calibrated surrogates $L$? 
In this paper we study \emph{polyhedral} surrogates, meaning piecewise-linear convex functions over prediction space $\reals^d$.
\citet{finocchiaro2019embedding} show that polyhedral surrogates are in a strong sense equivalent to an \emph{embedding}: % \emph{embedding} the discrete loss $\ell$ into $\reals^d$ by 
a mapping of each discrete prediction $r$ to some point $\varphi(r) \in \reals^d$, which optimizes $L$-loss if and only if $r$ optimizes $\ell$-loss.
While the authors show that every discrete loss $\ell$ can be so embedded, in the worst case the dimension required is $d=n-1$ where $n$ is the number of possible observations.
A surrogate dimension $d$ significantly below $n$, such as $O(\log n)$ for classification with an abstain option~\citep{ramaswamy2018consistent}, can lead to faster downstream optimization and computation, an effect that grows with $n$; this motivates us to understand when this dimension can be low.

@@

%In this work, we study \emph{embedding dimension}, which is the minimum dimension of input to a polyhedral surrogate that is calibrated with respect to the given discrete loss.
%Moreover, we use this tool to characterize \emph{how} embeddable a given discrete loss is, showing necessary conditions for embedding a given loss into $d$ dimensions.
In order to concisely discuss embedding dimension, we appeal to notation and terminology from the field of property elicitation \citep{savage1971elicitation, osband1985information-eliciting, gneiting2007strictly, lambert2009eliciting, lambert2018elicitation}, relating it to the language of calibrated surrogates as needed.

@@

\begin{definition}[Embedding dimension]
  We say a discrete loss $\ell: \R \to \nonnegreals^{\Y}$ is \emph{$d$-embeddable} if there exists a polyhedral surrogate $L: \reals^d \to \nonnegreals^{\Y}$ that embeds it.
  The \emph{embedding dimension} of $\ell$ is the smallest $d$ such that $\ell$ is $d$-embeddable.
\end{definition}

@@

We summarize the above results in the following theorem, together with one additional result, whose proof is in Appendix~\ref{app:1-dim}: if $\ell$ has any calibrated convex surrogate at all, it must have a polyhedral one.
(See Conjecture~\ref{conj:embed-equals-ccdim} regarding higher dimensional generalizations.)
\begin{theorem} \label{thm:1d-tfae}
  Let $\ell$ be a discrete loss eliciting a finite property $\gamma$.
  The following are equivalent: (1) $\gamma$ is orderable; (2) the intersection graph of $\gamma$ is a path; (3) the two conditions of Theorem \ref{thm:construct-1d-loss} are satisfied; (4) $\ell$ is $1$-embeddable; (5) $\ell$ has some polyhedral calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$; (6) $\ell$ has some convex calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
\end{theorem}
\jessiet{If I understand correctly, different orderings of the reports may yield different embeddings, and these different embeddings should elicit different properties.  How to we reconcile this in the theorem statement?  Or where am I wrong?}
\raft{I think you are right, but I don't see a contradiction (do you?)}

@@

Recall that if $\ell: \R \to \nonnegreals^{\Y}$ is embedded by $L: \reals^d \to \nonnegreals^{\Y}$, then each $r \in \R$ is embedded at some point $\varphi(r) \in \reals^d$.
In particular, $\varphi(r)$ must minimize $\inprod{p}{L(\cdot)}$ if and only if $r$ minimizes $\inprod{p}{\ell(\cdot)}$.
The key to our approach is to study as first-class objects the sets of all subgradients\footnote{Recall that a subgradient of e.g. the convex function $L(\cdot)_y: \reals^d \to \reals$ at a point $u$ is a vector $v \in \reals^d$ such that $L(u')_y \geq L(u)_y + \inprod{v}{u'-u}$ for all $u'$.} of $L$ at these embedding points.
The question of whether a calibrated polyhedral surrogate exists in $d$ dimensions essentially reduces to conditions on these sets alone.
\btw{The ``essentially'' is because we need loss values to match.  It's not clear if that comes for free, i.e., if you embed the correct property, can you always modify the loss to match $\ell$?}
In particular, we use the fact that a convex function is minimized at $u$ if and only if $\vec{0}$ is in its subgradient set at $u$.
Therefore, we consider collections of sets $T^r_y$, which intuitively aspire to be the subgradient sets of a calibrated polyhedral surrogate $L(\cdot)_y$ at $\varphi(r)$, denoted $\partial L(\varphi(r))_y$.
Throughout, we often take $r$ as implicit and suppress it from our notation for ease of exposition.
Note that if $L(\cdot)_y$ is a polyhedral function on $\reals^d$, then all of its subgradient sets are (bounded) closed polytopes~\citep{rockafellar1997convex}.

@@

  \vspace{-3.2em}
    \begin{align}
      V X^j                     &\leq B    & \text{(pointwise, $\forall j \in [l]$)}  \label{eqn:qp-constr-1} \\
      \sum_{y=1}^n p^j_y x^j_y &= \vec 0  & \text{($\forall j \in [l]$)}    \label{eqn:qp-constr-2}
    \end{align}
\end{definition}

@@

%% Bo: I think we can get away with no summary at the end.
%\paragraph{Conclusion.}
%\jessie{Not a fan, but a placeholder if we get too close to deadline.}
%In this paper, we present the notion of \emph{embedding dimension} in order to measure the efficiency of a surrogate loss function.
%While similar notions of surrogate efficiency have been studied before, we propose a characterization in $1$ dimension that generalizes to the construction of any calibrated convex surrogate.
%In higher dimensions, our characterization can only be applied to understanding the minimal dimension of a calibrated \emph{polyhedral} surrogate, although we conjecture that embedding dimension is equal to convex calibration dimension in general.
%In order to demonstrate our characterizations, we show an improved lower bound from previous work on the convex calibration dimension of $\abstain{\alpha}$, and a new lower bound on embedding dimension of $\abstain{1/2}$.
%
\newpage
% Acknowledgments---Will not appear in anonymized version
\acks{Jessie Finocchiaro was funded by NSF Graduate Research Fellowship under grant No.\ DGE 1650115.  This material is based upon work supported by the National Science Foundation under Grant No.\ 1657598.  We also thank Peter Bartlett, Nishant Mehta, and Hari Ramaswamy for helpful comments.  \raf{Others...?} \bo{Others...?} \jessie{Others...?}}
